Using traffic analysis to identify The Second Generation Onion Router

### 2011-04-03 thesis/thesis.tex
Bit of work on tor section, more detail, more detail on vulnerabilities

The goal of anonymity sometimes aligns with that of providing resistance to
censorship. Many censorship resistance systems are built on top of anonymity
providing designs, \textcite{Kopsell:2004:ABR:1029179.1029197} describes a
censorship resistant system built on top of AN.ON
\parencite{springerlink:10.1007/3-540-44702-4_7}. Publishing systems such as
Freedom \parencite{Goldberg:1999p2231} and Freenet \parencite{Clarke:2001p2435}
aim to provide censorship resistance as well as anonymity. Infranet uses
steganography to hide sensitive messages in innocuous content Infranet
\parencite{Feamster:2002p307}.

@@

In the Tor network, the intermediate nodes responsible for routing messages are
known as relays and are typically chained together to construct a circuit.
Clients connect to the Tor network by communicating with a special relay known
as a bridge, which constitutes the first relay in a circuit. Traffic is
repeatedly delivered through relays until it reaches another special relay known
as an exit node, where it is delivered unencrypted to the final destination.
Figure \ref{tor-network} shows this typical usage of the Tor network.

@@

Circuits are built incrementally, obtaining a session key from each
successive relay in the circuit and encrypting messages so that intermediate
relays are unable to decrypt the message. Since circuits can take some time to
build, they are pre-emptively created at regular intervals
\parencite[5]{Dingledine:2004p314}. Although Tor does not support many of the
traffic normalization techniques that provide strong anonymity, messages passed
through the Tor network are packaged in cells of a fixed size of 512 bytes
\parencite[5]{Dingledine:2004p314}. Tor also provides application level filters
that remove identifying features from common application protocols.

@@

\section{Tor Weaknesses}

Research into Tor weaknesses has included the use of hostile exit nodes to
capture private information \parencite{website:tor-password-leak}, attacks that
take advantages of web based technologies such as Javascript and Flash to reveal
user's identities \parencite{Abbott:2007p298} and a number of identity revealing
traffic analysis techniques
\parencite{Murdoch:2005p325,Abbott:2007p298,Evans:2009p315}.
\citeauthor{McLachlan:2009p197} reveals a method for identifying bridge
operators in \textcite{McLachlan:2009p197} using the Tor protocol directly.

@@

Tor employs strong encryption, the goal of which is to ensure that the data
transmitted is indistinguishable from random noise. To an observer, two
different streams of encrypted data look almost the same. Thus the traditional
means of identifying these protocols are ineffective.

This raises the question: is it possible to distinguish the traffic created by
these networks, from encrypted traffic produced by conventional software? This
question can be broken down further:

\begin{enumerate*}
  \item Can the traffic be classified using automated matching techniques?
  \item Do these networks have characteristics that make them readily
    distinguishable using heuristics based matching?
  \item Do they have characteristics that make them distinguishable using
    machine learning techniques?
  %\item How long does a user have to be
  %  connected to the network before a confident match can be made?
\end{enumerate*}

>>>>

More sophisticated circumvention tend to be based on anonymity providing
designs, \textcite{Kopsell:2004:ABR:1029179.1029197} describes a censorship
resistant system built on top of AN.ON
\parencite{springerlink:10.1007/3-540-44702-4_7}. Publishing systems such as
Freedom \parencite{Goldberg:1999p2231} and Freenet \parencite{Clarke:2001p2435}
aim to provide censorship resistance as well as anonymity. Infranet uses
steganography to hide sensitive messages in innocuous content Infranet
\parencite{Feamster:2002p307}.

@@

Traffic enters the Tor network through an onion proxy which accepts TCP streams.
Some identifying features are scrubbed from the data using application filters
before the data is relayed over the network through TLS \parencite{website:TLS}
encrypted connections. The intermediate nodes responsible for routing messages
are known as relays and are typically chained together to construct a circuit.
When traffic leaves the Tor network it is delivered by a special kind of relay
known as an exit node. At an exit node the data is transmitted in the original
format it was supplied to at the onion proxy.

Figure \ref{tor-network} shows the typical path a message takes through the Tor
network.

@@

The onion proxy builds circuits incrementally, first obtaining a session key
from each successive relay in a circuit. Once all session keys for a circuit
have been obtained, the message is broken into fixed sized cells of 512 bytes
and iteratively encrypted using the session key of each node in the circuit in
the reverse order that the data traverses the network. Cells come in two forms:
control cells and relay cells. Control cells are used to create and maintain
circuits, while relay cells contain commands for circuit maintenance and
additional data for verifying message integrity and identifying streams.

When the first node in the circuit receives a cell, it removes a layer by
decrypting the message using it's session key. The cell is delivered along the
circuit and at each node a layer is removed until the cell is completely
decrypted and can be delivered successfully. Since circuits can take some time
to build, they are pre-emptively created at regular intervals
\parencite[5]{Dingledine:2004p314}.

\section{Vulnerabilities and Weaknesses}

When designing a system such as Tor, a number of trade-offs have to be made
between the strength of the security provided and the convenience and
performance of the system. The Tor designers consciously prioritized low
latency, usability and flexibility against security goals such as security
against end to end attacks \parencite[4]{Dingledine:2004p314}. These design
decisions mean that Tor is vulnerable to several kinds of attacks, some of them
expected and some not.

A well known attack involves sniffing traffic that leaves exit nodes to
capture private information \parencite{website:tor-password-leak}. Many users
assume that Tor provides end to end encryption, and transmit private information
over the Tor network.

Most conventional attacks against secure networks are known as traffic analysis,
this is the process of examining information about the communications rather
than the information contained within them. This information may include the
size of messages communicated, their frequency and timing. Many researchers have
proposed traffic analysis techniques that allow attackers to reveal the
identities of Tor users.

Technologies such as Javascript and Flash can be embedded in web pages accessed
by Tor users, and have control over network resources. By injecting network
traffic with certain patterns alongside regular network traffic, Tor users can
be identified \parencite{Abbott:2007p298}.

Tor bridges, intended as a way to get censored users access to Tor are easily
identified. They are also vulnerable to clogging attacks which make bridge
operators more easily identified \textcite{McLachlan:2009p197}.

\textcite{Murdoch:2005p325} proposes a technique to identify users by estimating
the latency of individual Tor nodes using a hostile Tor node. The hostile Tor
node is able to send data to users using a predictable traffic pattern and
identify this pattern as it is repeated through the network, correlating streams
back to individuals.

This attack was shown to be impractical as the Tor network grew in size, with 
the increased number of users adding enough congestion to mask the introduced
identifying patterns \parencite{Evans:2009p315}. However a weakness in the Tor
design meant that particularly configured hostile nodes could amplify the
deliberately introduced congestion to make individuals identifiable on the
larger network.

This paper describes a simple traffic analysis technique which can be used to
distinguish Tor traffic from another form of commonly encrypted traffic.
Utilized properly this technique could be used to automatically identify Tor
nodes and block them, preventing Tor from being used as a censorship
circumvention tool.  Alternatively it could be used to isolate Tor users and
target them for attack, be it technically orientated such as a denial of service
attack or socially; coercion or arrest. Any direct attack on Tor users can
discourage participation, making the network less effective.

@@

\section{Traffic Analysis as Classification}

@@

\begin{enumerate*}
  \item Can Tor traffic be identified using automated matching techniques?
  \item Does Tor have characteristics that make it readily distinguishable using
    heuristics based matching?
  \item Do they have characteristics that make them distinguishable using
    machine learning techniques?
\end{enumerate*}
